\input{code/Afrac_SysID_values}
\input{code/Afrac_CLTF_values}
\input{code/SM_SysID_values}
\input{code/SM_CLTF_values}

\input{figures/DetCtrl_2D_35853}

\input{figures/Afrac_LoopStability}

\section{System identification and Controller Tuning}
\label{sec:sysid}

\input{figures/DetCtrl_2D_35854}

\input{figures/SM_LoopStability}

Before we attempted detachment control experiments, we took two system identification shots.
The data from the first system identification shot \#35853 is shown in Fig.\ref{fig:sysid_afrac}.
In this shot, we puffed in N$_2$ gas in steps of 1.0 V, 2.5V, and 4.0 V with puff duration of 1.5s each.
A corresponding response was seen in \Afrac but with a delay.
We later confirmed from post-shot EFIT data that the strike point was indeed within the real-time Langmuir Probe array and thus our \Afrac calculation was valid.
We fitted the measured data with a simple first-order plant model of gain K, time constant $\tau$, and time delay $L$ given by:
\textcolor{red}{This is just FOPDT like was used before? You might mention that this is the frequency space representation of the same FOPDT model that was used previously and cite past work where FOPDT was used for sys ID, like my ppcf 2022. Let them know this part didn't really change too much.}

\begin{equation}
 G(s) = \frac{K}{\tau s + 1}e^{-L s}
\label{eq:sysid}
\end{equation}

\textcolor{red}{define G and s}

The fit resulted in an identified model with \AfracK, \AfracTau, and \AfracL.
The fit is shown in Fig.\ref{fig:sysid_afrac}b.
Note that only the part of the time series data that was used in fit is shown for the fitted curve.
This fit was performed in the inter-shot interval during the experiment and has not been improved or modified after the experiment.

The controller gains were chosen by visualizing \ac{CLTF} of the system with chosen PI gains as shown in Fig.\ref{fig:cltf_afrac}.
Here, the frequency domain response of the plant model ($G(s)$) and PI controller ($T_{PI}(s)$) are plotted together.
When connected in series, this forms the \ac{OLTF} of the system ($O(s) = G(s) T_{PI}(s)$).
The frequency where \ac{OLTF} becomes 1.0 is called \ac{UGF}).
Phase margin is defined as the additional phase delay at \ac{UGF} that would make the system unstable by taking it to -180$^\circ$.
Additionally, we also define delay margin as the additional actuation delay that would make \ac{UGF} unstable.
The \ac{CLTF} is then calculated by solving the loop algebra in the Laplace domain:

\begin{equation}
 C(s) = \frac{G(s)}{1 + O(s)}
\label{eq:cltf}
\end{equation}

\input{figures/DetCtrl_2D_35857}

Because of the long delay, we chose to not use a derivative gain.
The goal of tuning was to push \ac{UGF} as high as possible (\AfracUGF) while keeping a reasonable phase margin (\AfracPhaseMargin) and margin for any additional actuation delay (\AfracDelayMargin).
This resulted in controller settings as: \AfracKp, \AfracTi, and \Afracstau, where K$_p$ is proportional gain, T$_i$ is integral time, and $\tau_s$ is pre-smoothing time constant.
This is still a very aggressive choice of controller, but provided that the system identification fit gave an unexpectedly high value of response time $\tau=1$s probably due to too much noise during small step inputs, we decided to go ahead with this controller choice.
The resulting PI controller transfer function is given by:

\begin{equation}
 T_{PI}(s) = K_p \left( \frac{1}{T_i s} + 1\right) \frac{1}{1 + \tau_s s}
\label{eq:PI}
\end{equation}

Unfortunately, the surrogate model was not configured properly in this system identification shot due to technical errors, so we repeated a system identification but this time we decided to keep the nitrogen valve in the constant open position, to look for any deviation in the behavior.
The data from this second system identification shot is shown in Fig.\ref{fig:sysid_sm}.
Despite all the limitations of DivControlNN as listed earlier, we still saw a good correlation in the DivControlNN heat flux output at the outer divertor with the injected gas as seen in Fig.\ref{fig:sysid_sm}b.
This is validated by estimated \Afrac in Fig.\ref{fig:sysid_sm}d showing an increase in detachment level as the predicted output heat flux decreases.
The strike point was maintained within the realtime Langmuir probe array (Fig.\ref{fig:sysid_sm}a) validating the output of \Afrac.
Note that we have not yet calibrated the DivControlNN model with any experimental data, so we treated the output as arbitrary units and later attempted to control the detachment with estimated changes to this arbitrary output.

We again fitted this system with a first-order system with delay as described in Eq.\ref{eq:sysid}.
The fit resulted in an identified model with \SMK, \SMTau, and \SML.
The fit is shown in Fig.\ref{fig:sysid_sm}b.
Here as well, the fitting shown was performed during the experiment in the inter-shot interval and has not been modified or optimized later.
The time domain in which the fitting curve is shown is the data where the system was fitted.
Admittedly, this fit was not very good and we did not believe the large lag value to be accurate.
% There is no physical reason for why the lag in the system would be higher when we use DivControlNN output as compared to \Afrac.
So for the purpose of tuning the controller, we arbitrarily set the system lag value to 100 ms.
The controller gains were chosen by visualizing \ac{CLTF} of the system with chosen PI gains as shown in Fig.\ref{fig:cltf_sm} and following the same procedure as we described for \Afrac controller tuning.
The resulting controller settings were: \SMKp, \SMTi, and \SMstau{} creating controller given by Eq.\ref{eq:PI}.
Here, we estimated to achieve a \ac{UGF} of \SMUGF, phase margin of \SMPhaseMargin, and delay margin of \SMDelayMargin.
This controller was also very aggressive, but we decided to go ahead with this controller choice given the limitations of the system identification fit and lack of time for further analysis in between the allotted run time of our experiment.

\textcolor{red}{Some notes to possibly include in discussion later: the sys ID for the two systems has pretty different times. DivControlNN is basically a density measurement plus a transformation, and core average density has additional transport timescales that should slow it down relative to a purely SOL response like seen in Isatmeasured. Afrac is a mix of both SOL and core timescales. So it makes sense if DivControlNN is slower, right? Not trusting this in the short time between shots might've gotten us in trouble when picking gains to go with DivCotnrolNN.}


\textcolor{red}{What you can also discuss somewhere (maybe here, maybe in discussion later) is that these controller tunings all are implicitly linearizing what we know to be a complicated and nonlinear system, and that, plus the presence of noise, means the probability of actual performance deviating from ideal predictions is pretty high and ad-hoc tuning adjustments after seeing the first closed loop test are common. So it was reasonable to try something to get more information; the first closed loop control test was a diagnostic.}
